---
title: "Lecture-16 Example Code"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook illustrates the code for lecture 16.

## Dependencies
This notebook requires only three additional package beyond base `R` packages:

```{r load-packages}
# tidyverse packages
library(dplyr)        # for pipe operator

# other packages
library(janitor)      # tables
library(testDriveR)   # data
```

## Load Data
This notebook requires the `auto17` data from `testDriveR`:

```{r load-data}
autoData <- auto17
```

## Part 1
### Questions 1 and 2.
First, we need to create an indicator for each vehicle that is `TRUE` if it is made by a German manufacturer. Then we want to subset our data so that we have only `id`, `german`, and `driveStr`.

```{r p1}
autoData %>% 
  mutate(german = ifelse(mfr == "BMW" | mfr == "Mercedes" | mfr == "Porsche" | 
                           mfr == "Volkswagen", TRUE, FALSE)) %>%
  select(id, german, driveStr) -> autoData
```

## Part 2
### Question 3
Next, we need to create a two-way contingency table of our data:

```{r p2}
autoData %>%
  tabyl(german, driveStr) %>%
  adorn_totals(where = c("row", "col")) %>%
  adorn_percentages(denominator = "row") %>%
  adorn_pct_formatting(digits = 3) %>%
  adorn_ns(position = "front")
```

We can see that a higher proportion of German vehicles are all wheel drive and rear wheel drive (owing to the Porsche sports cars).

## Part 3
### Question 4
We want to know if this variance that we observe visually is substantive, so we'll use a chi-square test:

```{r p3-q4}
(model <- chisq.test(autoData$german, autoData$driveStr))
```

The chi-square test ($\chi^{2} = 46.827, df = 4, p < .001$) indicates that there is substantial variation in drive train type based on whether or not the car is manufactured in Germany.

### Question 5
Even though we did not get a warning, we'll check to see whether we've violated the Cochran conditions:

```{r p3-q5}
model$expected
```

We can see no values less than 5 (though we come very close with one expected count), meaning we have not violated this assumption.

### Question 6
Had we violated Cochran conditions, we would have used Fisher's Exact Test to similaute a p-value:

```{r p3-q6}
fisher.test(autoData$german, autoData$driveStr, simulate.p.value = TRUE)
```

The Fisher's Exact test (p = .0005) indicates that there is substantial variation in drive train type based on whether or not the car is manufactured in Germany.
