---
title: "Lecture-16 Example Code"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: 
  github_document: default
  html_notebook: default 
---

## Introduction
This notebook illustrates the code for lecture 16.

## Dependencies
This notebook requires only three additional package beyond base `R` packages:

```{r load-packages}
# tidyverse packages
library(dplyr)        # for pipe operator
library(ggplot2)      # for mpg data

# other packages
library(janitor)      # tables
```

## Load Data
This notebook requires the `mpg` data from `ggplot2`:

```{r load-data}
autoData <- mpg
```

## Contingency Tables
### Basic Table
A basic contingency table is called using the `tabyl` function from `janitor`:

```{r contingency-basic}
autoData %>%
  tabyl(cyl, drv)
```

### Adding Totals
We can add row and column totals with an "adornment" - what `janitor` calls add-ons called after the initial `tabyl` function is called:

```{r contingency-totals}
autoData %>%
  tabyl(cyl, drv) %>%
  adorn_totals(where = c("row", "col"))
```

### Adding Percentages
Percentages can be used in place of frequencies by adding a second adornment:

```{r contingency-percents}
autoData %>%
  tabyl(cyl, drv) %>%
  adorn_totals(where = c("row", "col")) %>%
  adorn_percentages(denominator = "row")
```

### Formatting Percentages
Percentages can be formatting by adding another adornment:

```{r contingency-format-percents}
autoData %>%
  tabyl(cyl, drv) %>%
  adorn_totals(where = c("row", "col")) %>%
  adorn_percentages(denominator = "row") %>%
  adorn_pct_formatting(digits = 3)
```

### Adding Frequencies Back In
What if we want both frequencies and percentages? We can add them back in using a final adornment:

```{r contingency-frequencies}
autoData %>%
  tabyl(cyl, drv) %>%
  adorn_totals(where = c("row", "col")) %>%
  adorn_percentages(denominator = "row") %>%
  adorn_pct_formatting(digits = 3) %>%
  adorn_ns(position = "front")
```

## Calculating Chi-Squared
If we want to know if the differences observed between `cyl` and `drv` are substantive - would we see variation this extreme if there were no patterning - we can use the chi-square test. It is helpful to fit the model by wrapping the entire call in `()` so that we can get the model output *and* store the model in an object simultaneously. 

```{r chi-square}
(model <- chisq.test(mpg$cyl, mpg$drv))
```

The chi-square test ($\chi^{2} = 98.136, df = 6, p < .001$) indicates that there is substantial variation in cylinders by drive train type.

## Cochran conditions
Notice that a warning prints with our model output: 

> Chi-squared approximation may be incorrect

This is an indication that we have some issues with the model. *We should explore this assumption anyway even if there is no warning, however.* Our model object contains a matrix with expected values. We want to evaluate them:

```{r expected}
model$expected
```

There are three values here that are less than 5 and 3/12 = .25 - this is a sign we've ciolated the Cochran conditions. Furthermore, there is one value below 1, which is a significant concern.

We can neatly summarize this by creating a logical test:

```{r expected-lgl-5}
model$expected < 5
```

We can do the same for expected counts less than 1:

```{r expected-lgl-1}
model$expected < 1
```

## Fisher's Exact Test
In the case where we violate Cochran conditions, we can use Fisher's Exact Test to similaute a p-value:

```{r fisher}
fisher.test(mpg$cyl, mpg$drv, simulate.p.value = TRUE)
```

The Fisher's Exact test (p = .0005) indicates that there is substantial variation in cylinders by drive train type.
